---
title: "RepData_PeerAssessment1"
author: "Lester Burgwardt"
date: "11/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##  Activity Analysis
Read in the data into a data frame  

```{r Ingest data, echo = TRUE}
activity_file = 'C:/Users/Charlie/Documents/Coursera/DataScience/ReproducibleResearch/Week2/RepData_PeerAssessment1/repdata_data_activity/activity.csv'
activity_df = read.csv(activity_file)
```

Take a look at the first three lines  

```{r}
head(activity_df,3)
```
To calculate the total number of steps taken per day:  
1. Group by date  
2. Then find sum for each date  
  
```{r}
totalSteps = aggregate(activity_df$steps, by=list(Date=activity_df$date), FUN=sum)
hist(totalSteps$x,
     main='Histogram of Total Steps',
     xlab='Total Steps',
     ylab='Frequency')

#Insert a page break after the plot
cat("  \n")

```
The mean and median of the total number of steps taken per day are:  
```{r}
mean(totalSteps$x, na.rm=TRUE)
median(totalSteps$x, na.rm=TRUE)
```
Plot the average number of steps per interval and show which interval has the highest average number of steps.  
```{r}
averageStepsPerInterval = aggregate(activity_df$steps, by=list(Interval=activity_df$interval), FUN=mean, na.rm=TRUE)
plot(averageStepsPerInterval$Interval, averageStepsPerInterval$x,
     type = "l",
     main='Time Series of Average Steps per Time Interval',
     xlab='Time Interval',
     ylab='Number of Steps')
averageStepsPerInterval$Interval[which.max(averageStepsPerInterval$x)]
```
Number of missing values:  
```{r}
sum(is.na(activity_df$steps))
```
Set missing values to median of 5-minute interval.  
The new dataset is imputed_df.

```{r}
medianStepsPerInterval = aggregate(activity_df$steps, 
                                   by=list(Interval=activity_df$interval), 
                                   FUN=median, na.rm=TRUE)
imputed_df <- activity_df
for(i in 1:length(imputed_df$steps)){
        if (is.na(imputed_df$steps[i])){
                imputed_df$steps[i] <- 
                        medianStepsPerInterval$x[which(
                                medianStepsPerInterval$Interval == imputed_df$interval[i])]
        }
        
}
totalImputedSteps = aggregate(imputed_df$steps, by=list(Date=imputed_df$date), FUN=sum)
hist(totalImputedSteps$x,
     main='Histgram of Total Steps Including Imputed Values',
     xlab='Total Steps',
     ylab='Frequency')
#Insert a page break after the plot
cat("  \n")

```
The mean and median of the total number of steps including imputed values taken per day are:  
```{r}
mean(totalImputedSteps$x, na.rm=TRUE)
median(totalImputedSteps$x, na.rm=TRUE)
```
The mean changed from 10766.19 to 9503.869 steps per day,  
the median changed from 10765 to 10395 steps per day.  

Set up the day type variable as a factor with weekday or weekend  
```{r}
library(chron)
imputed_df$daytype = is.weekend(as.Date(imputed_df$date))
imputed_df$daytype[imputed_df$daytype] <- 'weekend'
imputed_df$daytype[!as.logical(imputed_df$daytype)] <- 'weekday'
imputed_df$daytype <- as.factor(imputed_df$daytype)

```
Make a two panel time series for weekdays and weekends  
```{r}

averageImputedStepsWeekday = aggregate(
        imputed_df$steps[which(imputed_df$daytype=='weekday')],
        by=list(Date=imputed_df$date[which(imputed_df$daytype=='weekday')]),
        FUN=mean)
averageImputedStepsWeekend = aggregate(
        imputed_df$steps[which(imputed_df$daytype=='weekend')],
        by=list(Date=imputed_df$date[which(imputed_df$daytype=='weekend')]),
        FUN=mean)

par(mfrow=c(2,1))
par(cex = 0.6)
par(mar = c(0, 0, 0, 0), oma = c(4, 7, 0.5, 0.5))
par(tcl = -0.25)
plot(as.Date(averageImputedStepsWeekend$Date),averageImputedStepsWeekend$x, 
     type='l')
plot(as.Date(averageImputedStepsWeekday$Date),averageImputedStepsWeekday$x, 
     type='l')

mtext("Date", side = 1, outer = TRUE, cex = 0.7, line = 2.2,col = "grey20")
mtext("Number of Steps", side = 2, outer = TRUE, cex = 0.7, line = 2.2,col = "grey20")

```


